A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines

### 2019-07-29 latex/paper.tex
Edited discussion

\subsection{Effect of Network File System}
We used an NFS as it is a common approach to store a large amount of data. However,
it limits the available bandwidth since all workers interact with one disk. This
causes data-intensive applications to take a significant amount of time for IO. As we
expect, this effect is amplified when too many tasks simultaneously try to operate
with the NFS at the same time as
Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
shows. A parallel file system, such as Lustre~\cite{lustre}, would likely improve
IO times of parallel applications simultaneously writing to shared storage.

\subsection{Effect of data-intensive applications}
In our experiments, Incrementation, Histogram (NumPy) and Histogram for Spark have
low compute time while BIDS App example and Histogram for Dask have significantly
longer compute time. When adding additional workers, we observe that the highly
data-intensive applications scale poorly while compute-intensive applications scale
almost linearly This is due to the limitation from the NFS, as mentioned above.

Data transfer accounts for the majority of the execution time for data-intensive
applications (see
Fig.~\ref{fig:inc_gantt},\ref{fig:histo_spark_gantt},\ref{fig:histo_np_gantt}). We
see that when a low number of IO tasks are overlapping, this diminishes the bandwidth
limitation from the NFS. We want to explore solutions to desynchronize IO tasks in
the aim to reduce the makespan of data-intensive applications.

\subsection{Effect of the Global Interpreter Lock}
Python's GIL prevents multi-threading in Python. Dask tries to minimize the
consequences of the GIL by using external Python extensions. We found that Dask has
issues to parallelize Python functions while Spark did not (see
Fig.~\ref{fig:histo_tt_worker}). However, Dask was easily able to parallelize
application with functions in Numpy (see
Fig.~\ref{fig:inc_tt_worker},\ref{fig:histo_np_tt_worker}) and command-line
applications (see Fig.~\ref{fig:bids_tt_worker}). This is because, for Dask, the
instances are configured to have 1 worker process with 8 threads. Since the Histogram
implementation does not release the GIL it can only use one of the thread at a given
time thus it slows down the computation significantly.

One should be careful to configure the workers properly depending on the type
of functions executed in an application.

\subsection{Effect of tasks' size}
We measured the effect of various block sizes for constant BigBrain image for
Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
results in more IO compensated by lower overhead time. On the other hand, smaller
block size results in higher overhead time compensated by reduced IO time.

The number of partitions also has an impact on the size of the tasks. We set a
partition for each block or subject. There is a trade-off for the number of
partitions. On the one hand, a low amount of partition results in larger tasks thus
in lower scheduling overhead. On the other hand, a higher number of partition
results in smaller tasks hence more tasks fit in the workers' memory insuring that
they always have something to work on.

Varying the block size is equivalent to varying the size of the partitions. We
observed no difference when varying the task size while using an NFS. However, if
one uses a parallel shared file system, we suppose it would be preferable to use a
larger block size as the IO time should be reduced substantially. It is worth noting
that the Dask developers suggest, in
\href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
use a partition size such that workers have 2-3 times as many chunks available to
work on. More experiments would have to be performed to gain insight on an optimal
number of partitions to use for a given set of data-intensive applications.

\subsection{Other comments}
Spark performance should be reduced by the serialization of Python functions to Java,
however, our results show no significant impact of it.

Previous work showed that Dask has a the largest start-up overhead~\cite{Mehta:17}
when processing few subjects. Our Gantt charts show that Spark has a higher start-up
overhead. We think this is due to the larger amount of subjects we process.

The same study showed that Dask has a lower end-to-end runtime than Spark in a
neuroimaging analysis application. In our results, there is no significant difference
between the engines. We think this is because the applications we chose are highly
data-intensive while the one used in their study is compute intensive.

\section{Conclusion}
We presented a comparison of two Big Data engines, Apache Spark and Dask.
We studied the engines on three data-intensive neuroimaging applications
representative of common use cases. Our results show no substantial
performance difference between the engines, except for compute-intensive
Python applications where Dask is slower than Spark due to parallelism loss
coming from Python's Global Interpreter Lock. Interestingly, differences in
engine overheads do not impact performance due to their impact on data
transfer times: higher overheads are almost exactly compensated by a lower
transfer times when data transfers saturate the bandwidth. These results
suggest that future research should focus on strategies to reduce the
impact of data transfers on applications.

>>>>

\subsection{Comparison with previous studies}

Overall, our experiments did not show any substantial performance
difference between Spark and Dask. In contradiction with the experiments
of~\cite{Mehta:17}, we did not observe consequential overhead differences
due to serialization or startup costs. This apparent contradiction is most
likely coming from differences in applications: while the applications in
~\cite{Mehta:17} were mostly compute-intensive, ours were more data-intensive.
Consequently, in our case, potential
overhead differences were compensated by reverse trends in data transfer
time. Since our focus are Big Data applications, this leads us to conclude that 
engines are equivalent. However, this conclusion doesn't invalidate the results of~\cite{Mehta:17} that 
would apply to more compute-intensive applications.

\subsection{Overhead / data transfer compensation}

The observed almost exact compensation between differences in engine
overheads and in data transfer times (see for instance
Figure~\ref{fig:inc_tt_worker}) is not surprising. Indeed, when the transfer
bandwidth is saturated, and in absence of service thrashing,
desynchronizing file transfers usually does not reduce the makespan of data
transfers. To take an extreme example, the makespan of $n$ concurrent file
transfers of size $F$ on a system of bandwidth $B$ would be $nF/B$, as $n$
transfers would share the bandwidth, which equals to the time required
to transfer the $n$ files sequentially. In other words, saturated bandwidths
give room for extra engine overhead as long as it desynchronizes data
transfers. This could be an interesting idea to explore in data transfer optimization.

\subsection{Effect of the shared file system}

We used NFS as our shared file system as it is a common and
easy-to-configure solution to share data between compute nodes. The
performance of NFS, however, degrades substantially when accessed by
concurrent tasks, as shown in the Gantt charts in
Figures~\ref{fig:inc_gantt}, ~\ref{fig:histo_gantt}, ~\ref{fig:histo_np_gantt} 
and~\ref{fig:bids_gantt}. This is due to the limitations imposed by network and disk
bandwidths. While network bandwidth can hardly be overcome, disk bandwidth
can be increased by storing data to multiple disks addressed by a parallel
file system such a Lustre~\cite{lustre}. Such a configuration would likely
reduce the impact of data transfers, although the main observed patterns
should remain.

\subsection{Effect of Python's Global Interpreter Lock}

Although Dask reduces the impact of Python's GIL on parallelism by using external
Python extensions, our Python histogram experiment (Figure
~\ref{fig:histo_tt_worker}) showed that issues remain in Dask to
parallelize Python functions. However, Dask was easily able to parallelize
the same application when implemented with Numpy
(Figure~\ref{fig:histo_np_tt_worker}) because in this case the computation
happened in C, outside of the GIL. Similarly, the computation in
command-line applications happens in a sub-process, hence outside of the
GIL, which explains why the BIDS apps example application was not impacted
by this issue. Although implemented in Python, Incrementation was not
impacted either, as computation was only emulated through sleep time.

This behavior is also the consequence of our configuration of Dask that
created 1 worker process with 8 threads. Since the Histogram Python
implementation does not release the GIL it can only use one of the thread
at a given time thus it slows down the computation significantly. This
behavior highlights the need to configure the workers properly depending on
the type of functions executed in an application.

\subsection{Effect of data partitioning}

The number of data partitions also has an impact on application
performance. In our experiments, we set a partition for each block or
subject, resulting in a maximum of 750 partitions for the Big Brain
applications, and 1,397 partitions for the BIDS app. The number of
partitions obeys a classical trade-off: small numbers of partitions reduce 
overhead, while large numbers of
partitions increase parallelism. Since both Spark and Dask can
work with custom partition sizes, we do not expect substantial differences
coming from variations of this parameter.

% Varying the block size is equivalent to varying the size of the partitions. We
% observed no difference when varying the task size while using an NFS. However, if
% one uses a parallel shared file system, we suppose it would be preferable to use a
% larger block size as the IO time should be reduced substantially. It is worth noting
% that the Dask developers suggest, in
% \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
% use a partition size such that workers have 2-3 times as many chunks available to
% work on. More experiments would have to be performed to gain insight on an optimal
% number of partitions to use for a given set of data-intensive applications.

\section{Conclusion}
We presented a comparison of two Big Data engines, Apache Spark and Dask.
We studied the engines on three data-intensive neuroimaging applications
representative of common use cases. Overall, our results show no substantial
performance difference between the engines. Interestingly, differences in
engine overheads do not impact performance due to their impact on data
transfer times: higher overheads are almost exactly compensated by a lower
transfer times when data transfers saturate the bandwidth. These results
suggest that future research should focus on strategies to reduce the
impact of data transfers on applications.
